### 2014-04-25 tech-project/led_trans.tex

% An LED Demo Translation
\section{LED Demo}

\begin{figure}
\centering
\begin{verbatim}
/*
 * Generated by the RESOLVE to C translator. 
 * This file should not be modified.
 */
#ifndef __LEDS_TEST_H
#define __LEDS_TEST_H

#include ".../RESOLVE.h"
#include ".../Facilities/.../Std_Boolean_Fac.h"
#include ".../Facilities/.../Std_Integer_Fac.h"
#include ".../Facilities/.../Std_Clock_Fac.h"
#include ".../Concepts/.../Leds_Template.h"
#include ".../Concepts/.../Std_Leds_Realiz.h"
#include ".../Concepts/.../Toggling_Realiz.h"

typedef struct Led_Facility Led_Facility;
struct Led_Facility {
    Leds_Template* core;
    Toggling_Capability_for_Leds_Template* 
                      Toggling_Capability;
};
Led_Facility Led_Facility_Var;
void Leds_Test_create();
void Leds_Test_destroy();

#endif
\end{verbatim}
\caption{LED\_Demo.h}
\end{figure}

\begin{figure}
\centering
\begin{verbatim}
/*
 * Generated by the RESOLVE to C translator. 
 * This file should not be modified.
 */
#include "Leds_Test.h"
void Leds_Test_create() {
  Std_Boolean_Fac_create();
  Std_Integer_Fac_create();
  Std_Clock_Fac_create();
  r_type_ptr __arg_0 = Std_Integer_Fac_Var.core->
  createFromInteger(4, Std_Integer_Fac_Var.core->Integer);
  Led_Facility_Var.core = 
  new_Std_Leds_Realiz_for_Leds_Template(__arg_0);
  Led_Facility_Var.Toggling_Capability = 
  new_Toggling_Realiz_for_Toggling_Capability_of_LED_Template(
								      Led_Facility_Var.core);
  Std_Integer_Fac_Var.core->Integer->
  destroy(__arg_0, Std_Integer_Fac_Var.core->Integer);
}

void Leds_Test_destroy() {
    free_Std_Leds_Realiz_for_Leds_Template(
    				    Led_Facility_Var.core);
    Std_Boolean_Fac_destroy();
    Std_Integer_Fac_destroy();
}
\end{verbatim}
\caption{LED\_Test.c}
\end{figure}