### 2014-04-25 tech-project/led_trans.tex

% An LED Demo Translation
\section{LED Demo}

\begin{figure*}
\begin{minipage}{0.4\textwidth}
\begin{verbatim}
/*
 * Generated by the RESOLVE to C translator. 
 * This file should not be modified.
 */
#ifndef __LED_TELOS_DEMO_H
#define __LED_TELOS_DEMO_H

#include ".../RESOLVE.h"
#include ".../Facilities/.../Std_Bool_Fac.h"
#include ".../Facilities/.../Std_Int_Fac.h"
#include ".../Facilities/.../Std_Clock_Fac.h"
#include ".../Concepts/.../LED_Template.h"
#include ".../Concepts/.../Std_LED_Realiz.h"
#include ".../Concepts/.../Toggling_Realiz.h"

typedef struct LED_Facility LED_Facility;
struct LED_Facility {
    LED_Template* core;
    Toggling_Capability_for_LED_Template* 
                      Toggling_Capability;
};
LED_Facility LED_Fac_Var;

void LED_Telos_Demo_create();
void LED_Telos_Demo_destroy();
#endif
\end{verbatim}
\end{minipage}
\hspace{1.5cm} 
\begin{minipage}{0.4\textwidth}
\begin{verbatim}
/*
 * Generated by the RESOLVE to C translator. 
 * This file should not be modified.
 */
#include "LED_Telos_Demo.h"

void LED_Telos_Demo_create() {
  Std_Bool_Fac_create();
  Std_Int_Fac_create();
  Std_Clock_Fac_create();
  r_type_ptr __arg_0 = Std_Int_Fac_Var.core->
    createFromInt(4, Std_Int_Fac_Var.core->Integer);
  LED_Fac_Var.core = 
    new_Std_LED_Realiz_for_LED_Template(__arg_0);
  LED_Fac_Var.Toggling_Capability = 
    new_Toggling_Realiz_for_
    Toggling_Capability_of_LED_
    Template(LED_Fac_Var.core);
  Std_Int_Fac_Var.core->Integer->
    destroy(__arg_0, Std_Int_Fac_Var.core->Integer);
}

void LED_Telos_Demo_destroy() {
    free_Std_LED_Realiz_for_LED_Template(
    				    LED_Fac_Var.core);
    Std_Bool_Fac_destroy();
    Std_Int_Fac_destroy();
}
        ...
\end{verbatim}
\end{minipage}
\end{figure*}