### 2014-05-03 tech-project/telos_platform.tex
fuck

\section{The Telos Platform}

The Telos mote\cite{polastre:2005} is a programmable, low power, wireless sensing device developed at UC Berkeley. Its hardware includes an MSP430-F1611 microcontroller containing 10kB of RAM and 48kB of programmable flash memory\cite{msp:2011}. A CC2420 radio stack provides the mote with broadcast and receiving capabilities, while optional sensing capabilities may be added in the form of light, humidity, and temperature sensors. The mote also contains three onboard LEDs: One blue, one red, and one green.